---
title: "Cost-Effectiveness Analysis: Tirzepatide vs Semaglutide"
author: "Keerthana Choutha"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This report evaluates the cost-effectiveness of **Tirzepatide** versus **Semaglutide** for treating Type 2 Diabetes Mellitus (T2DM). It includes:

- ICER calculation using QALYs
- One-way sensitivity analysis
- Export of results to Excel

# Load Required Libraries

```{r load-packages}
install.packages(c("dplyr", "writexl"))  # Run only once
library(dplyr)
library(writexl)
```

# Define Base Inputs

```{r base-inputs}
base_values <- list(
  mace_tp_t = 0.425,
  mace_tp_s = 0.444,
  mort_tp_t = 0.512,
  mort_tp_s = 0.367,
  cvd_cost = 14888.32,
  dm_cost = 13107.60,
  drug_cost_t = 1079.77,
  drug_cost_s = 997.59
)

utility_no_mace <- 0.85
utility_mace <- 0.70
utility_dead <- 0
```

# ICER Function (Based on QALYs)

```{r icer-function}
calc_icer_qaly <- function(params) {
  p_mace_t <- params$mace_tp_t
  p_mort_t <- params$mort_tp_t
  p_no_mace_t <- max(0, 1 - p_mace_t - p_mort_t)

  cost_t <- (p_mace_t * params$cvd_cost) +
            (p_no_mace_t * params$dm_cost) +
            params$drug_cost_t

  qaly_t <- (p_mace_t * utility_mace) +
            (p_no_mace_t * utility_no_mace) +
            (p_mort_t * utility_dead)

  p_mace_s <- params$mace_tp_s
  p_mort_s <- params$mort_tp_s
  p_no_mace_s <- max(0, 1 - p_mace_s - p_mort_s)

  cost_s <- (p_mace_s * params$cvd_cost) +
            (p_no_mace_s * params$dm_cost) +
            params$drug_cost_s

  qaly_s <- (p_mace_s * utility_mace) +
            (p_no_mace_s * utility_no_mace) +
            (p_mort_s * utility_dead)

  delta_cost <- cost_s - cost_t
  delta_qaly <- qaly_s - qaly_t
  icer <- delta_cost / delta_qaly

  return(list(
    icer = icer,
    cost_t = cost_t,
    cost_s = cost_s,
    qaly_t = qaly_t,
    qaly_s = qaly_s
  ))
}
```

# Sensitivity Ranges (±20%)

```{r sensitivity-ranges}
sensitivity <- data.frame(
  parameter = names(base_values),
  base = unlist(base_values)
) %>%
  mutate(
    low = base * 0.8,
    high = base * 1.2
  )
```

# One-Way Sensitivity Analysis

```{r one-way-sensitivity}
icer_results <- data.frame(Parameter = character(), Low = numeric(), High = numeric())

for (i in 1:nrow(sensitivity)) {
  param <- sensitivity$parameter[i]

  low_params <- base_values
  low_params[[param]] <- sensitivity$low[i]
  low_result <- calc_icer_qaly(low_params)

  high_params <- base_values
  high_params[[param]] <- sensitivity$high[i]
  high_result <- calc_icer_qaly(high_params)

  icer_results <- rbind(icer_results, data.frame(
    Parameter = param,
    Low = low_result$icer,
    High = high_result$icer
  ))
}

icer_results_clean <- icer_results %>%
  mutate(
    Low = round(Low, 2),
    High = round(High, 2),
    Range = round(abs(High - Low), 2)
  )
```

# Base Case Summary

```{r base-summary}
base_result <- calc_icer_qaly(base_values)

base_summary <- data.frame(
  Group = c("Tirzepatide", "Semaglutide"),
  Cost = round(c(base_result$cost_t, base_result$cost_s), 2),
  QALY = round(c(base_result$qaly_t, base_result$qaly_s), 4)
)

base_icer <- data.frame(ICER = round(base_result$icer, 2))
```

# Export to Excel

```{r export-excel}
write_xlsx(list(
  "One-Way Sensitivity" = icer_results_clean,
  "Base ICER" = base_icer,
  "Cost and QALY Summary" = base_summary
), "icer_sensitivity_results.xlsx")
```

# Results Summary

```{r show-results}
icer_results_clean
```

```{r show-base}
base_summary
```

```{r show-icer}
base_icer
